
## 4.6.2

```{r}
set.seed(2020)
A <- sample(50:80, 10, replace=T)
print(A)
pick <- A > 60 # is asking 
whichMetTheCondition <- which(pick)
```

```{r}
A[pick]
A[witchMetTheCondition]
```

```{r}
grades <- c(51,70,79,78,67,74,80)
# mean student grades
```

```{r}
puck <- grades > 70 # or
witchIs70plus <- witch(pick)
```

```{r}
grades[pick]
```

  * `...[pick]` : retrieve those whose grades >70
  
  * `grades...`: retreieve from grades for those whose grades > 70
  
  * `grades[pick]` retrieve from grades (`grades...`) for those whose grades >70 (`...[pick]`)
  

```{r}
grades <- c(51,70,79,78,67,74,80)
# same 7 student's gender
gender <- c('female','female','male','male','female','male','male')
```
```{r}
pick <- grades > 70
gender[pick]
```

```{r}
gender[pick] # means
```

  * retrieve from **gender**
  
  * for those whose **genders > 70**
  

```{r}
dataSet1 <- 
  data.frame(
    school_id = c("001", "002", "003", "004"),
    course1_grade=c(55, 83, 92, 73),
    course2_grade=c(50, 88, 72, 77)
  )
```

For those who have a grade >=80 in course1, what are their grades in course2?

```{r}
# For those who have a grade >=80 in course1
pick <- c(55, 83, 92, 73) >= 80
pick <- dataSet1$course1_grade >= 80

# what are their grades in course2
# retrieve their grades in course2
c(50, 88, 72, 77)[pick]
dataSet1$course2_grade[pick]
```

```{r}
pick <- dataSet1$course1_grade >= 80
```
  * For those whose **course1 grades** are >=80,

```{r}
daraSet1$course2_grade[pick] #[pick] conditional retrieval
```

  * their **course2 grades** are 

***

For those who have course1 grade >=80,  what are their records

  * A record is a collect of features from the same observation (student).

```{r}
dataSet1 |> names( ) # |> only R 4.10
names(dataSet1)

pick <- dataSet1$course1_grade >
```
```{r}
dataSet1$school_id[pick]
dataSet1$school1_grade[pick]
dataSet1$school2_grade[pick]
```

```{r}
class(dataSet1)
```

  * `[.row, .col]`
  
```{r}
view(dataSet1)
```

```{r}
dataSet1[pick, ] -> subsample
View(subsample)
```

  * For those whose course1 grades >= 80 (`[pick, ]`),
  
  * their records (`dataSet1`) are: dataSet1[pick, ]`
  
***
Conditional retrieval.

  * For those whose ...,
  
  * retrieve xxx.

```{r}
xxx[...]
```

```{r}
subset(dataSet1, course1_grade >=80) -> subsample2
View(subsample2)
# same as
pick <- dataSet1$course1_grade >= 80
dataSet1[pick, ]
```

```{r}
pick <- c(F, T, NA, T)

c(1, 2, 3, 4)[pick]
```


# 4.6.3

```{r}
gender = c('Female','Male','Male', 'Female')
age = c(28,41,42,33)
residence = c('South','South','North', 'North')
color=c("yellow", "pink", "blue", "green")
```

  * For those who...
  
   *are male

```{r}
pick <- gender == "Male"
```

  * For those who...
  
   * are male; (AND/OR)
   
   *lives in the south

```{r}
# 1.
pick_male <- gender == "Male"
# not 1.
!pick_male
# 2.
pick_south <- residence == "south"
```

```{r}
# 1. AND 2.
pick_male & pick_south
# 1. OR 2.
pick_male | pick_south
```

***

```{r}
pick <- list()
```

```{r}
# For those who are "Male": 
pick$male <- gender == "Male"
print(pick$male)
gender
# For those who lives in the "South:
pick$south <- residence == "South"
print(pick$south)
residence
```

```{r}
# For those who are "Male" 
#   AND
#   lives in the "South",
pick$male_south <- 
  pick$male & pick$south

print(pick$male_south)

# what are their ages?
age[pick$male_south]
```

For those who are “Male” AND lives in the “South”
(`[pick$male_south]`), what are their ages `age...`?


***

For those who are “Male” OR lives in the “South”

```{r}
pick$maleOsouth <-
  pick$male | pick$south
```

what are their ages?

```{r}
gender
residence
ages
age[pick$maleOsouth]
```

***

For those who are male or from south, but not both

  * exclude `pick$male & pick$south`
  
```{r}
pick$maleXOsouth <-
  xor(pick$male, pick$south) -> pick$maleXOsouth

gender
residence
age
age[pick$maleXOsouth]
```

```{r}
pick$male
gender
color
color[pick$male]
color[pick$south]
color[pick$male & pick&south]
color[pick$male | pick$south]
color[xor(pick$male,pick$south)]
```

```{r}
# For those who are male, but not from South
pick$maleXsouth <- 
  pick$male & !pick$south
color[pick$maleXsouth]

# For those who are neither male nor from south
pick$XmaleXsouth <-
  !pick$male & !pick$south
color[pick$XmaleXsouth]
```

```{r}
!(! pick$male) # the same as
pick$male
```

```{r}
!(pick$male & pick$south) # the same as
!pick$male | !pick$south
```

```{r}
!(pick$male | pick$south) # the same as
!pick$male & !pick$south
```

```{r}
any(pick$male)
all(pick$male)
```

```{r}
pick2 <- c(T, T, NA)
any(pick2) # T
all(pick2) # NA
amy(na.omit(pick2))
all(na.omit(pick2))
```

```{r}
pick3 <- c(F, F, NA)
any(pick3) # NA
all(pick3) # T
```



